Abstractions for Type-Level Programming

### 2022-04-05 preprint.tex
Copy paste more of the dependent scala  related work

Proving termination or totality of programs is a non-goal of our system.
Unlike proof assistants, Scala programs do not manipulate proof terms.
Consequently, the lack of totality checks does not affect Scala's present notion of safety.
Exceptions or infinite loops in the evaluation of a `dependent` method would prevent the completion of type-checking.

@@

% TODO: Compare to pure subtype systems

>>>>

Proving termination or totality of programs is a non-goal of our system.
Unlike proof assistants, Scala programs do not manipulate proof terms.
Consequently, the lack of totality checks does not affect Scala's present notion of safety.
Exceptions or infinite loops in the evaluation of a |dependent| method would prevent the completion of type-checking.

@@

Dependently-typed lambda calculi with subtyping were described at least as far back as 1988~\citep{cardelli1988structural}.
His type system is much more expressive than ours and allows bounded quantification over both types and terms using the notion of a |Type| type and power types.
Unlike our system, which is designed with the concrete evaluation of types in mind, Cardelli does not provide a semantics for his system and leaves the equivalence relation among types unspecified.

In~\citep{aspinall1994subtyping} Aspinall introduces a dependently-typed system with subtyping \emph{and} singleton types that resembles ours in its type language.
His equivalence relation on types is more powerful and is not syntax-directed, unlike our type evaluation relation.
Furthermore, singleton types in his work are indexed by the type through which equality is "viewed", thereby enabling a form of polymorphism beyond ours.
Aspinall's system also has primitives types and allows for atomic subtyping among them, but no congruence rules, so partially-widened forms like |(_: Any) :: Nil| cannot be represented.

Pure subtype systems~\citep{hutchins2010pure} are different in that they only contain a single syntactic category for both terms and types, and a single relation, subtyping, that subsumes typing, subtyping and type evaluation of our system.
However, their system allows for partially-widened types similar to ours and also enables the computation with different levels of precision.
For instance, it is able to conclude that |Int + 5| can be approximated as |Int|.

Dependent-object types~\citep{amin2017type} model the core of Scala's type system and include type members and path-dependent types, which are not represented in our formalism.
Though they introduce a form of dependency, path-dependent types were not designed for type-level computation, rendering their goals largely orthogonal to ours.
